Mergebiomark167With10X

### 2021-10-05 MergeBiomarkWith10X/MergeBiomark167With10X.Rmd
Show that anchor integration is the best method for biomark/10x integration

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/MergeBiomarkWith10X/")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
```

# Set global variables
```{r}
projectName <- "Mergebiomark167With10X"
jackstraw.dim <- 40
```

```{r}
source("msAggr_AnalysisCode/read_10XGenomics_data.R")
source("msAggr_AnalysisCode/PercentVariance.R")
source("msAggr_AnalysisCode/Mouse2Human_idconversion.R")
```

@@

```{r}
color.palette <- c(
	"coral",
	"chartreuse4",
	"goldenrod1",
	"cadetblue1",
	"burlywood",
	"brown",
	"brown1",
	"blue",
	"blue4",
	"azure3",
	"aquamarine",
	"antiquewhite",
	"cadetblue",
	"gold3",
	"black",
	"darkgreen",
	"deeppink",
	"darkviolet",
	"darkturquoise",
	"darkslategray",
	"darksalmon",
	"darkorchid1",
	"darkolivegreen2",
	"forestgreen",
	"dodgerblue",
	"green",
	"lightpink",
	"lightcoral",
	"khaki1",
	"maroon",
	"peru",
	"lightseagreen",
	"lightsalmon",
	"plum",
	"moccasin",
	"tan",
	"tan1", 
	"red", 
	"purple",
	"khaki4",
	"black", 
	"plum4"
)
```

@@

Will need to include:
cell list
probe list
expression matrix

```{r}
biomark.order <- read.table("origBiomark_cellIDorder.txt", sep = "\t", header = FALSE, row.names = 1)
biomark.order["clust.ID"] <- NA
biomark.order[1:49, "clust.ID"] <- 0
biomark.order[50:87, "clust.ID"] <- 17
biomark.order[88:102, "clust.ID"] <- 10
biomark.order[103:117, "clust.ID"] <- 0
biomark.order[118:127, "clust.ID"] <- 3
biomark.order[128:150, "clust.ID"] <- 0
biomark.order[151:189, "clust.ID"] <- 11
biomark.order[190:281, "clust.ID"] <- 0
```

```{r}
# move fluorophor markers to metadata column, remove for expression dataset
fluor.ids <- colnames(biomark.data[grep("pos", colnames(biomark.data))])
fluor.table <- biomark.data[fluor.ids]
biomark.data.T <- biomark.data.T[!(rownames(biomark.data.T) %in% fluor.ids),]
```

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(biomark.data.T)){
	ngenes <- length(which(!(biomark.data.T[colnames(biomark.data.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}

# Match ids from original hierarchical clustering assignments to cells retained in Biomark167
i1 <- match(row.names(biomark.meta), row.names(biomark.order))
biomark.meta[names(biomark.order)[1]] <- lapply(biomark.order[1], `[`, i1)

# match normalized fluorescene 
i1 <- match(row.names(biomark.meta), row.names(fluor.table))
biomark.meta[names(fluor.table)[1:ncol(fluor.table)]] <- lapply(fluor.table[1:ncol(fluor.table)], `[`, i1)

@@

@@

```{r warning=FALSE}
setwd("../cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
```

@@

## merge objects
```{r}
seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
seurat.object@meta.data$orig.ident <- as.factor(seurat.object@meta.data$orig.ident)
seurat.object@meta.data$pid <- as.factor(seurat.object@meta.data$pid)
seurat.object@meta.data$wid <- as.factor(seurat.object@meta.data$wid)
seurat.object
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
```

@@

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("orig.ident", "nFeature_RNA", "pid", "nCount_RNA"))
```

### Save merged object
```{r}
# saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

@@

```{r}
clust3.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster3cellIDs.txt", col.names = "clust03")
clust3.cells <- sapply(clust3.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust17.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster17cellIDs.txt", col.names = "clust17")
clust17.cells <- sapply(clust17.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust10.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster10cellIDs.txt", col.names = "clust10")
clust10.cells <- sapply(clust10.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
clust11.cells <- read.table(file = "/Users/heustonef/Desktop/10XGenomicsData/msAggr_scRNASeq/Seuratv1_clusterCellIDs/cluster11cellIDs.txt", col.names = "clust11")
clust11.cells <- sapply(clust11.cells, function(x) paste0(gsub("CMP", "CMPm2", x), "-1"))
```

## map cluster ids from Seuratv1
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
seurat.object@meta.data$clust.ID <- as.factor(seurat.object@meta.data$clust.ID)
```

```{r fig.width=15, fig.height=5}
DimPlot(seurat.object, split.by = "clust.ID", group.by = "orig.ident", pt.size = 2)
```

## Split object for anchor-based integration

```{r}
cmp.subset <- SplitObject(seurat.object, split.by = "orig.ident")$`10X`
DimPlot(cmp.subset, group.by = "clust.ID")
```

## Follow integration protocol

https://satijalab.org/seurat/archive/v3.0/integration.html

```{r}
tech.list <- SplitObject(seurat.object, split.by = "orig.ident")

for(i in 1:length(tech.list)){
	tech.list[[i]] <- NormalizeData(tech.list[[i]], verbose = FALSE)
	tech.list[[i]] <- FindVariableFeatures(tech.list[[i]], selection.method = "vst", nfeatures = 10, verbose = FALSE)
}
```

## identify anchors
```{r}
# reference.list <- tech.list["10X"]
tech.anchors <- FindIntegrationAnchors(object.list = tech.list, dims = 1:30, anchor.features = all.genes, k.anchor = 20, reduction = 'rpca', k.filter = NA)
# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 5)
```
## tech.integrated PCA

@@

```{r fig.height = 5, fig.width=15}
DimPlot(tech.integrated, split.by = "clust.ID", shape.by = "orig.ident", pt.size = 2, cols = c("black", "red"), reduction = "pca")
```

```{r fig.height=10, fig.width=10}
png(filename = "tech.integrated_biomark167Fluor.png", height = 1600, width = 1600)
FeaturePlot(seurat.object, features = c("Cd34pos", "Cd135pos", "Cd9pos", "Cd36pos", "Cd41pos", "Cd48pos", "Cd123pos", "Cd150pos", "Cd135pos"), split.by = "clust.ID", reduction = "pca", cols = c("gray92", "blue"), keep.scale = "all", ncol = 1)
dev.off()

>>>>

# Updates

## 20211005
Compared to archive 1, this version imports Biomark Ct spreadsheets that do not include fluorophore data

```{r setup}
knitr::opts_knit$set(root.dir = "~/Desktop/10XGenomicsData/msAggr_scRNASeq/MergeBiomarkWith10X/")
```

```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
```

# Set global variables
```{r}
projectName <- "Mergebiomark167With10X"
```

```{r}
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/ColorPalette.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/PercentVariance.R")
source("~/Desktop/10XGenomicsData/msAggr_scRNASeq/RFunctions/read_10XGenomics_data.R")
```

@@

@@

```{r}
biomark.data <- read.table("/Users/heustonef/Desktop/10XGenomicsData/BioMark/Biomark_CT.txt", sep = "\t", header = TRUE, row.names = 1)
biomark.data[is.na(biomark.data)] <- 0

# Limit to CMP data
biomark.data <- biomark.data[grepl("CMP", rownames(biomark.data)) == TRUE,]
dim(biomark.data)
```
Convert to pseudoexpression data
```{r}
# generate list of reference Ct averages, calculating average for ref genes where ref_gene Ct > 0
ref_list <- c("Actb", "B2m", "Kit")
biomark.data["refCt"] <- NA

for(i in 1:nrow(biomark.data)){
	ref_cts <- c()
	for(j in ref_list){
		if(biomark.data[i, j] > 0){
			ref_cts <- c(ref_cts, biomark.data[i, j])
		}
	}
	biomark.data[i, "refCt"] <- mean(ref_cts)
}

# generate dCt col
dbiomark.data <- biomark.data[, 1:ncol(biomark.data) -1]
dbiomark.data[dbiomark.data == 0] <- NA
dbiomark.data <- dbiomark.data - biomark.data[, "refCt"]

# create pseudo expression matrix
pseudo.biomark <- 2^-dbiomark.data

@@

```{r}
# fix non-MGI gene names
colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'Cd117'] <- 'Kit'
colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'Cd123'] <- 'Il3ra'
colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'H2.Aa'] <- 'H2-Aa'
colnames(pseudo.biomark)[colnames(pseudo.biomark) == 'H2.Ab1'] <- 'H2-Ab1'
pseudo.biomark <- subset(pseudo.biomark, select = -Kng2)
```

```{r warning=FALSE, message=FALSE}
biomark.meta <- data.frame(stringr::str_split_fixed(rownames(pseudo.biomark), "_", 3)[,2:3], row.names = rownames(pseudo.biomark))
colnames(biomark.meta) <- c("wid", "pid")

pseudo.biomark.T <- data.table::transpose(pseudo.biomark)
colnames(pseudo.biomark.T) <- rownames(pseudo.biomark)
rownames(pseudo.biomark.T) <- colnames(pseudo.biomark)

biomark.meta['nFeature_RNA'] <- NA
for(cell in colnames(pseudo.biomark.T)){
	ngenes <- length(which(!(pseudo.biomark.T[colnames(pseudo.biomark.T > 0) == cell])))
	biomark.meta[rownames(biomark.meta) == cell, 'nFeature_RNA'] <- ngenes
}

```{r}
pseudo.biomark.T[is.na(pseudo.biomark.T)] <- 0
pseudo.biomark.T <- as.sparse(pseudo.biomark.T)
biomark <- CreateSeuratObject(pseudo.biomark.T, project = "Biomark", assay = "RNA", meta.data = biomark.meta)
```

@@

```{r warning=FALSE}
setwd("~/Desktop/10XGenomicsData/cellRanger/") # temporarily changing wd only works if you run the entire chunk at once
data_file.list <- read_10XGenomics_data(sample.list = "CMPm2")
object.data <-Read10X(data_file.list)
cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = "10X")
```

@@

## merge objects
```{r}
seurat.object <- merge(cmp.object, biomark, project = 'mergebiomark')
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMPm2'] <- '10X'
seurat.object@meta.data$orig.ident[seurat.object@meta.data$orig.ident == 'CMP'] <- 'Biomark'
# seurat.object@meta.data[is.na(seurat.object@meta.data)] <- 0
# seurat.object@meta.data$orig.ident <- factor(seurat.object@meta.data$orig.ident)
# seurat.object@meta.data$pid <- factor(seurat.object@meta.data$pid)
# seurat.object@meta.data$wid <- factor(seurat.object@meta.data$wid)
seurat.object
```

# Traditional normalize and scaling approaches
trying sctransform and normalizeData/scaledata
```{r}
# seurat.object <- SCTransform(seurat.object, vars.to.regress = c("wid"))
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 10)
```

@@

```{r}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("wid", "pid", "orig.ident"))
```

### Save merged object
```{r}
saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```
## Run PCA
```{r}
seurat.object <- RunPCA(seurat.object, ndims.print = 1:5, nfeatures.print = 5, features = all.genes)
```

@@

# scRNASeq integration	
 neither sctransform nor normalizedata/scaledata gave what looked like reasonable results.

https://satijalab.org/seurat/articles/integration_introduction.html

```{r}
tech.list <- SplitObject(seurat.object, split.by = "orig.ident")
tech.list <- lapply(X = tech.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 10)
})
```

## identify anchors
```{r}

tech.anchors <- FindIntegrationAnchors(object.list = tech.list, anchor.features = all.genes, k.anchor = 20, k.filter = NA)
# tech.integrated <- IntegrateData(anchorset = tech.anchors, dims = 1:30, k.weight = 20, normalization.method = "SCT")
tech.integrated <- IntegrateData(anchorset = tech.anchors, k.weight = 5)
```
## tech.integrated PCA

@@
